---
title: "Pest introduction risk based on international trade"
author: "Aaron Plex"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Organizing trade data
```{r, message=FALSE}
library(tidyverse)

cropTradeWTO<-read.csv("WTO-Tomato-Trade-2011-2020.csv")
cropTradeWTO<-cropTradeWTO[,c("Reporting.Economy", "Partner.Economy", "Year", "Value")]
colnames(cropTradeWTO)<-c("ImportCountry", "ExportCountry", "Year", "Value")
years<-length(unique(cropTradeWTO$Year))

cropTradeWTO <- cropTradeWTO %>%
  group_by(ExportCountry, ImportCountry) %>%
  summarise(YearTradeMean = sum(Value, na.rm = TRUE)/years) %>%
  filter(YearTradeMean > 0) %>%
  mutate(YearTradeMean = (YearTradeMean / max(YearTradeMean))^(1/4))

#Replacing long names to abbreviations in edgelist
cropTradeWTO[cropTradeWTO == c("Antigua and Barbuda")] <- c("ANTI")
cropTradeWTO[cropTradeWTO == c("Bahrain, Kingdom of")] <- c("Bahrain")
cropTradeWTO[cropTradeWTO == c("Bolivia, Plurinational State of")] <- c("Bolivia")
cropTradeWTO[cropTradeWTO == c("Bosnia and Herzegovina")] <- c("BiH")
cropTradeWTO[cropTradeWTO == c("Central African Republic")] <- c("CF")
cropTradeWTO[cropTradeWTO == c("China, mainland")] <- c("China")
cropTradeWTO[cropTradeWTO == c("Democratic Republic of the Congo")] <- c("Congo-Kinshasa")
cropTradeWTO[cropTradeWTO == c("Congo, Democratic Republic of the")] <- c("Congo-Kinshasa")
cropTradeWTO[cropTradeWTO == c("Congo, Republic of the")] <- c("Congo Republic")
cropTradeWTO[cropTradeWTO == c("Czech Republic")] <- c("Czechia")
cropTradeWTO[cropTradeWTO == c("Dominican Republic")] <- c("DOM")
cropTradeWTO[cropTradeWTO == c("Hong Kong, China")] <- c("Hong Kong")
cropTradeWTO[cropTradeWTO == c("Iran (Islamic Republic of)")] <- c("Iran")
cropTradeWTO[cropTradeWTO == c("C\xf4te d'Ivoire")] <- c("Ivory Coast")
cropTradeWTO[cropTradeWTO == c("Saint Kitts and Nevis")] <- c("KN")
cropTradeWTO[cropTradeWTO == c("Kuwait, the State of")] <- c("Kuwait")
cropTradeWTO[cropTradeWTO == c("Kyrgyzstan")] <- c("Kyr") 	
cropTradeWTO[cropTradeWTO == c("Kyrgyz Republic")] <- c("Kyr")
cropTradeWTO[cropTradeWTO == c("Lao People's Democratic Republic")] <- c("Laos") 
cropTradeWTO[cropTradeWTO == c("Lebanese Republic")] <- c("Lebanon")
cropTradeWTO[cropTradeWTO == c("Macao, China")] <- c("Macao")
cropTradeWTO[cropTradeWTO == c("Micronesia (Federated States of)")] <- c("Micronesia")
cropTradeWTO[cropTradeWTO == c("Moldova, Republic of")] <- c("Moldova")
cropTradeWTO[cropTradeWTO == c("Democratic People's Republic of Korea")] <- c("North Korea")
cropTradeWTO[cropTradeWTO == c("Papua New Guinea")] <- c("PNG")
cropTradeWTO[cropTradeWTO == c("Russian Federation")] <- c("Russia")
cropTradeWTO[cropTradeWTO == c("Saudi Arabia, Kingdom of")] <- c("SB") 
cropTradeWTO[cropTradeWTO == c("Slovak Republic")] <- c("Slovakia")
cropTradeWTO[cropTradeWTO == c("Korea, Republic of")] <- c("South Korea")
cropTradeWTO[cropTradeWTO == c("Sao Tome and Principe")] <- c("STD")
cropTradeWTO[cropTradeWTO == c("Syrian Arab Republic")] <- c("Syria")
cropTradeWTO[cropTradeWTO == c("China, Taiwan Province of")] <- c("Taiwan")
cropTradeWTO[cropTradeWTO == c("United Republic of Tanzania")] <- c("Tanzania")
cropTradeWTO[cropTradeWTO == c("Trinidad and Tobago")] <- c("TT")
cropTradeWTO[cropTradeWTO == c("T\xfcrkiye")] <- c("Turkey")
cropTradeWTO[cropTradeWTO == c("United Arab Emirates")] <- c("UAE")
cropTradeWTO[cropTradeWTO == c("United Kingdom")] <- c("UK")
cropTradeWTO[cropTradeWTO == c("United States of America")] <- c("USA")
cropTradeWTO[cropTradeWTO == c("Saint Vincent and the Grenadines")] <- c("VC")
cropTradeWTO[cropTradeWTO == c("Venezuela, Bolivarian Republic of")] <- c("Venezuela")
cropTradeWTO[cropTradeWTO == c("Viet Nam")] <- c("Vietnam")

cropTradeWTO <- data.frame(cropTradeWTO)

# Generating a country node list
CountryNodes1 <- as.vector(cropTradeWTO$ExportCountry)
CountryNodes2 <- as.vector(cropTradeWTO$ImportCountry)
CountryNodes <- rbind(CountryNodes1, CountryNodes2)
CountryNodes <- as.vector(CountryNodes)
CountryNodes <- as.matrix(CountryNodes)
colnames(CountryNodes) <- c("ExportCountry")
CountryNodes <- unique(CountryNodes)
```


## Host distribution at the country level
```{r}
host.distribution<-read.csv("FAOSTAT-Tomato-harvested-area-2011-2020.csv")
host.distribution<-host.distribution[,c("Area","Year","Value")]
colnames(host.distribution)<-c("ExportCountry","Year","HostArea")

host.distribution <- as_tibble(host.distribution) %>%
  group_by(ExportCountry) %>% # Calculating mean annual harvested area
  summarise(HostArea = mean(HostArea, na.rm = TRUE)) %>%
  filter(HostArea > 0) %>%  # Keeping country where crop is present
  mutate(HostArea = log10(HostArea+11/10)) # Log-transformation to reduce skenness

host.distribution[host.distribution == c("Antigua and Barbuda")] <- c("ANTI")
host.distribution[host.distribution == c("Bahrain, Kingdom of")] <- c("Bahrain")
host.distribution[host.distribution == c("Bolivia (Plurinational State of)")] <- c("Bolivia")
host.distribution[host.distribution == c("Bosnia and Herzegovina")] <- c("BiH")
host.distribution[host.distribution == c("Central African Republic")] <- c("CF")
host.distribution[host.distribution == c("China, mainland")] <- c("China")
host.distribution[host.distribution == c("Democratic Republic of the Congo")] <- c("Congo-Kinshasa")
host.distribution[host.distribution == c("Congo, Democratic Republic of the")] <- c("Congo-Kinshasa")
host.distribution[host.distribution == c("Congo, Republic of the")] <- c("Congo Republic")
host.distribution[host.distribution == c("Congo")] <- c("Congo Republic")
host.distribution[host.distribution == c("Dominican Republic")] <- c("DOM")
host.distribution[host.distribution == c("China, Hong Kong SAR")] <- c("Hong Kong")
host.distribution[host.distribution == c("Iran (Islamic Republic of)")] <- c("Iran")
host.distribution[host.distribution == c("C\xf4te d'Ivoire")] <- c("Ivory Coast")
host.distribution[host.distribution == c("Côte d'Ivoire")] <- c("Ivory Coast")
host.distribution[host.distribution == c("Saint Kitts and Nevis")] <- c("KN")
host.distribution[host.distribution == c("Kyrgyzstan")] <- c("Kyr")
host.distribution[host.distribution == c("Lao People's Democratic Republic")] <- c("Laos")
host.distribution[host.distribution == c("China, Macao SAR")] <- c("Macao")
host.distribution[host.distribution == c("Micronesia (Federated States of)")] <- c("Micronesia")
host.distribution[host.distribution == c("Republic of Moldova")] <- c("Moldova")
host.distribution[host.distribution == c("Netherlands (Kingdom of the)")] <- c("Netherlands")
host.distribution[host.distribution == c("Democratic People's Republic of Korea")] <- c("North Korea")
host.distribution[host.distribution == c("Papua New Guinea")] <- c("PNG")
host.distribution[host.distribution == c("Russian Federation")] <- c("Russia")
host.distribution[host.distribution == c("Saudi Arabia")] <- c("SB")
host.distribution[host.distribution == c("Republic of Korea")] <- c("South Korea")
host.distribution[host.distribution == c("Sao Tome and Principe")] <- c("STD")
host.distribution[host.distribution == c("Syrian Arab Republic")] <- c("Syria")
host.distribution[host.distribution == c("China, Taiwan Province of")] <- c("Taiwan")
host.distribution[host.distribution == c("United Republic of Tanzania")] <- c("Tanzania")
host.distribution[host.distribution == c("Trinidad and Tobago")] <- c("TT")
host.distribution[host.distribution == c("Türkiye")] <- c("Turkey")
host.distribution[host.distribution == c("United Arab Emirates")] <- c("UAE")
host.distribution[host.distribution == c("United Kingdom of Great Britain and Northern Ireland")] <- c("UK")
host.distribution[host.distribution == c("United Kingdom")] <- c("UK")
host.distribution[host.distribution == c("United States of America")] <- c("USA")
host.distribution[host.distribution == c("Saint Vincent and the Grenadines")] <- c("VC")
host.distribution[host.distribution == c("Venezuela (Bolivarian Republic of)")] <- c("Venezuela")
host.distribution[host.distribution == c("Viet Nam")] <- c("Vietnam")
```


## Pest distribution at the country level
```{r message=FALSE, warning=FALSE}
PestDistribution <- read.csv("Tuta absoluta.csv")
PestDistribution<-PestDistribution[,c("Region","Presence", "Extent", "First.record","USOrder")]
colnames(PestDistribution)<-c("ExportCountry","Presence", "Extent", "First.record","USOrder")
PestDistribution<-PestDistribution[PestDistribution$ExportCountry!="",]

# Identifying which regions are sub national observations
is.subnational = (PestDistribution[grepl(" - ", PestDistribution$ExportCountry), ])
subnations = is.subnational$ExportCountry

# Excluding pest observations at sub national extent
PestDistribution = PestDistribution %>%
  filter(!ExportCountry %in% subnations)

# Replacing extent categories with values in a numeric scale between 1 to 5
PestDistribution<-PestDistribution %>%
  mutate(Extent = ifelse(Extent == "-"  | Extent == "Extent not recorded" & Presence == "Present", "Present", Extent))
PestDistribution$ExtentCat<-PestDistribution$Extent

PestDistribution[PestDistribution$Extent == "Widespread",3] <- c("5")
PestDistribution[PestDistribution$Extent == "Common",3] <- c("5")
PestDistribution[PestDistribution$Extent == "Native",3] <- c("5")
PestDistribution[PestDistribution$Extent == "Present",3] <- c("4")
PestDistribution[PestDistribution$Extent == "Localized",3] <- c("3")
PestDistribution[PestDistribution$Extent == "Exporting interceptions",3] <- c("3")
PestDistribution[PestDistribution$Extent == "Few occurrences",3] <- c("2")
PestDistribution[PestDistribution$Extent == "Sporadic",3] <- c("2")
PestDistribution[PestDistribution$Extent == "Transient under surveillance",3] <- c("2")
PestDistribution[PestDistribution$Extent == "Transient under eradication",3] <- c("1")
PestDistribution[PestDistribution$Extent == "Single occurrence",3] <- c("1")
PestDistribution[PestDistribution$Extent == "No longer present",3] <- c("0")
PestDistribution[PestDistribution$Extent == "Intercepted only",3] <- c("0")
PestDistribution[PestDistribution$Extent == "Eradicated",3] <- c("0")
PestDistribution[PestDistribution$Extent == "Unreported",3] <- c("0")
PestDistribution$Extent<-as.numeric(PestDistribution$Extent)

PestDistribution$USOrder<-ifelse(PestDistribution$USOrder=="yes","yes","no")
PestDistribution$Biosecurity<-ifelse(PestDistribution$USOrder=="yes",10,1)

# Replacing long country names
PestDistribution[PestDistribution == c("Antigua and Barbuda")] <- c("ANTI")
PestDistribution[PestDistribution == c("Bahrain, Kingdom of")] <- c("Bahrain")
PestDistribution[PestDistribution == c("Bolivia (Plurinational State of)")] <- c("Bolivia")
PestDistribution[PestDistribution == c("Bosnia and Herzegovina")] <- c("BiH")
PestDistribution[PestDistribution == c("Central African Republic")] <- c("CF")
PestDistribution[PestDistribution == c("China, mainland")] <- c("China")
PestDistribution[PestDistribution == c("Democratic Republic of the Congo")] <- c("Congo-Kinshasa")
PestDistribution[PestDistribution == c("Congo, Democratic Republic of the")] <- c("Congo-Kinshasa")
PestDistribution[PestDistribution == c("Congo, Republic of the")] <- c("Congo Republic")
PestDistribution[PestDistribution == c("Czech Republic")] <- c("Czechia")
PestDistribution[PestDistribution == c("Dominican Republic")] <- c("DOM")
PestDistribution[PestDistribution == c("China, Hong Kong SAR")] <- c("Hong Kong")
PestDistribution[PestDistribution == c("Iran (Islamic Republic of)")] <- c("Iran")
PestDistribution[PestDistribution == c("C\xf4te d'Ivoire")] <- c("Ivory Coast")
PestDistribution[PestDistribution == c("Saint Kitts and Nevis")] <- c("KN")
PestDistribution[PestDistribution == c("Kyrgyzstan")] <- c("Kyr")
PestDistribution[PestDistribution == c("Lao People's Democratic Republic")] <- c("Laos")
PestDistribution[PestDistribution == c("China, Macao SAR")] <- c("Macao")
PestDistribution[PestDistribution == c("Micronesia (Federated States of)")] <- c("Micronesia")
PestDistribution[PestDistribution == c("Republic of Moldova")] <- c("Moldova")
PestDistribution[PestDistribution == c("Democratic People's Republic of Korea")] <- c("North Korea")
PestDistribution[PestDistribution == c("Papua New Guinea")] <- c("PNG")
PestDistribution[PestDistribution == c("Russian Federation")] <- c("Russia")
PestDistribution[PestDistribution == c("Saudi Arabia")] <- c("SB")
PestDistribution[PestDistribution== c("Republic of Korea")] <- c("South Korea")
PestDistribution[PestDistribution == c("Sao Tome and Principe")] <- c("STD")
PestDistribution[PestDistribution == c("Syrian Arab Republic")] <- c("Syria")
PestDistribution[PestDistribution == c("China, Taiwan Province of")] <- c("Taiwan")
PestDistribution[PestDistribution == c("United Republic of Tanzania")] <- c("Tanzania")
PestDistribution[PestDistribution == c("Trinidad and Tobago")] <- c("TT")
PestDistribution[PestDistribution == c("United Arab Emirates")] <- c("UAE")
PestDistribution[PestDistribution == c("United Kingdom")] <- c("UK")
PestDistribution[PestDistribution == c("United Kingdom of Great Britain and Northern Ireland")] <- c("UK")
PestDistribution[PestDistribution == c("United States")] <- c("USA")
PestDistribution[PestDistribution == c("United States of America")] <- c("USA")
PestDistribution[PestDistribution == c("Saint Vincent and the Grenadines")] <- c("VC")
PestDistribution[PestDistribution == c("Venezuela (Bolivarian Republic of)")] <- c("Venezuela")
PestDistribution[PestDistribution == c("Viet Nam")] <- c("Vietnam")

PestHostdf<-merge(PestDistribution, host.distribution, by = 'ExportCountry', all = TRUE)

analysis.year<-2024
TradingCountries = merge(PestHostdf, CountryNodes, by = 'ExportCountry', all.y = TRUE)
TradingCountries = TradingCountries %>%
  mutate(First.record = as.numeric(First.record)) %>%
  distinct(ExportCountry, .keep_all = TRUE) %>%
  replace_na(list(Presence = "Unreported", 
             Extent = 0,
             First.record = analysis.year,
             ExtentCat = "Unreported",
             HostArea = log10(11/10))) %>%
  mutate(pest.age = analysis.year+1-First.record,
         country.type = ifelse(HostArea==log10(11/10),"Intermediary","Producer"))
```


## Converting trade matrix into a pest movement network
```{r, message=FALSE}
library(igraph)
TradeNet <- graph_from_data_frame(d= cropTradeWTO, 
                                  vertices = TradingCountries, 
                                  directed = T)

#Extracting an international trade matrix
tradeMatrix = as.matrix(as_adjacency_matrix(TradeNet, attr = "YearTradeMean"))

#Creating a matrix of pest extent
pestmatr = TradingCountries$Extent 
pestmatr1 <- matrix(pestmatr, , 1) #source country
pestmatr2 <- matrix(pestmatr, 1, ) #destination country
pestmatrix <- pestmatr1 %*% (1/(pestmatr2+1))
pestmatrix <- (2^pestmatrix)/(2^5)
pestmatrix=as.matrix(pestmatrix)

columnnames=c(TradingCountries$ExportCountry)
colnames(pestmatrix)=c(columnnames)
rownames(pestmatrix)=c(columnnames)

#Creating a matrix of pest duration in a country
pest.duration<-TradingCountries$pest.age
pest.age1 <- matrix(pest.duration, ,1) #source country
pest.age2 <- matrix(pest.duration, 1, )
pest.age2[] <- 1
page.matrix <- pest.age1 %*% pest.age2
#page.matrix <- page.matrix / max(page.matrix)

colnames(page.matrix)=c(columnnames)
rownames(page.matrix)=c(columnnames)

#Creating a matrix of host availability
hostmatr = TradingCountries$HostArea
hostmatr = hostmatr / max(hostmatr)
hostmatr1 <- matrix(hostmatr, , 1)
hostmatr2 <- matrix(hostmatr, 1, )
hostmatrix <- hostmatr1 %*% hostmatr2
hostmatrix=as.matrix(hostmatrix)

colnames(hostmatrix)=c(columnnames)
rownames(hostmatrix)=c(columnnames)

##Creating a matrix of pest-specific biosecurity
biomatr = TradingCountries$Biosecurity # yes = 2, no = 1
biomatr = 1 / biomatr
biomatr1 <- matrix(biomatr, , 1)
biomatr2 <- matrix(biomatr, 1, )
biomatrix <- biomatr1 %*% biomatr2
biomatrix=as.matrix(biomatrix)

colnames(biomatrix)=c(columnnames)
rownames(biomatrix)=c(columnnames)

#Gravity model for pest trade movement potential which is 
#trade matrix time pest matrix times host matrix
#I propose the name of this combination:
#Geographic trade risk for pest introduction
geoTradeRisk = pestmatrix * tradeMatrix * hostmatrix * page.matrix * biomatrix

#To convert from matrix to an edgelist
library(reshape2)
PestMoveLinks=melt(geoTradeRisk) %>%
  rename(ExportCountry=Var1,
         ImportCountry=Var2,
         GeoTradeRisk=value) %>%
  mutate(GeoTradeRisk=replace(GeoTradeRisk, GeoTradeRisk == 0.0000000, 0)) %>%
  filter(GeoTradeRisk>0)

#Generating a network for international pest movement
pestMoveNet = graph_from_data_frame(d = PestMoveLinks, 
                                  vertices = TradingCountries, 
                                  directed = T)
```


## A joint measure of pest introduction risk
```{r, fig.width=18, fig.height=15, dpi=200, message=FALSE}
library(ggraph)
library(RColorBrewer)

# Calculating the relative risk of pest introduction in a country from any other country in the world
No<-dim(geoTradeRisk)
joint.risk<-rep(NaN, No[1])
maxi<-max(geoTradeRisk)

for (i in 1:No[1]) {
  joint.risk[i] <- 1-prod(1-geoTradeRisk[1:No[1],i]/maxi)
}

V(pestMoveNet)$ExtentCat<-factor(V(pestMoveNet)$ExtentCat,
                                 levels = c("Unreported", "Eradicated", "No longer present", "Intercepted only", "Transient under eradication", "Transient under surveillance", "Few occurrences", "Exporting interceptions", "Localized", "Present", "Widespread"))

joint.net <- ggraph(pestMoveNet, layout = "centrality", cent = joint.risk^(1/4))+
  geom_edge_link(aes(edge_width = GeoTradeRisk, 
                      start_cap = circle(.3), 
                      end_cap = circle(.3)),                  
                  edge_colour = "darkgrey",
                  arrow = arrow(angle = 30, 
                                length = unit(0.15, "inches")))+
  geom_node_point(aes(fill = ExtentCat,
                      color = ExtentCat,
                      size = HostArea), 
                  shape = 21)+
  geom_node_text(aes(label = name), size = 6, 
                 repel = TRUE)+
  scale_edge_width_continuous(range = c(0.005,1))+
  scale_size_continuous(range = c(1,10))+
  scale_fill_manual(values = c("Unreported" = "#FCB519FF", "Eradicated" = "#FCB519FF", "No longer present" = "#FCB519FF", "Intercepted only" = "#FCB519FF", "Transient under eradication" = "#F98C0AFF", "Transient under surveillance" = "#F98C0AFF", "Few occurrences" = "#E35932FF", "Exporting interceptions" = "#BB3754FF", "Localized" = "#BB3754FF", "Present" = "#89226AFF", "Widespread" =  "#56106EFF"))+
  scale_color_manual(values = c("Unreported" = "#FCB519FF", "Eradicated" = "#FCB519FF", "No longer present" = "#FCB519FF", "Intercepted only" = "#FCB519FF", "Transient under eradication" = "#F98C0AFF", "Transient under surveillance" = "#F98C0AFF", "Few occurrences" = "#E35932FF", "Exporting interceptions" = "#BB3754FF", "Localized" = "#BB3754FF", "Present" = "#89226AFF", "Widespread" =  "#56106EFF"))+
  coord_fixed()+
  theme_graph()+
  theme(legend.position = "right", 
        legend.key.size = unit(1, 'cm'))

joint.net + 
  guides(fill = guide_legend(title = "Reported distribution: \nTuta absoluta",
                                title.position = "top", 
                                override.aes = list(size = 10), 
                             title.theme = element_text(size = 15), 
                             label.theme = element_text(size = 15)),
         color = guide_legend(title = "Reported distribution: \nTuta absoluta",
                                title.position = "top", 
                                override.aes = list(size = 10), 
                             title.theme = element_text(size = 15), 
                             label.theme = element_text(size = 15)),
            size = guide_legend(title = "log10( \ntomato \nacreage)", 
                                title.theme = element_text(size = 15), 
                                label.theme = element_text(size = 15)), 
            edge_width = guide_legend(title = "Pest trade \nmovement potential", 
                                      title.theme = element_text(size = 15), 
                                      label.theme = element_text(size = 15))
         )
```

## Introduction risk in the Caribbean Region and USA
```{r, fig.width=15, fig.height=11, dpi=250, message=FALSE}
focused.countries<-c("ANTI","Bahamas","Barbados","Cuba","DOM","Dominica","Haiti",
                     "Jamaica","KN","Saint Lucia","TT","VC","USA")
focused.PestMove<-PestMoveLinks %>%
  filter(ImportCountry %in% focused.countries)
focused.TradingCountries<-TradingCountries %>% 
  filter(ExportCountry %in% unique(c(focused.PestMove$ExportCountry,focused.PestMove$ImportCountry)))
focused.PestNet = graph_from_data_frame(d = focused.PestMove,
                                        vertices = focused.TradingCountries,
                                        directed = T)
V(focused.PestNet)$ExtentCat<-factor(V(focused.PestNet)$ExtentCat,
                                 levels = c("Unreported", "Eradicated", "No longer present", "Intercepted only", "Transient under eradication", "Transient under surveillance", "Few occurrences", "Exporting interceptions", "Localized", "Present", "Widespread"))

focused.net <- ggraph(focused.PestNet, layout = "stress")+
  geom_edge_link(aes(edge_width = GeoTradeRisk, 
                      start_cap = circle(.3), 
                      end_cap = circle(.3)),                  
                  edge_colour = "darkgrey",
                  arrow = arrow(angle = 30, 
                                length = unit(0.15, "inches")))+
  geom_node_point(aes(fill = ExtentCat,
                      color = ExtentCat,
                      size = HostArea), 
                  shape = 21)+
  geom_node_text(aes(label = name), size = 6, 
                 repel = TRUE)+
  scale_edge_width_continuous(range = c(0.3,2.5))+
  scale_size_continuous(range = c(1,10))+
  scale_fill_manual(values = c("Unreported" = "#FCB519FF", "Eradicated" = "#FCB519FF", "No longer present" = "#FCB519FF", "Intercepted only" = "#FCB519FF", "Transient under eradication" = "#F98C0AFF", "Transient under surveillance" = "#F98C0AFF", "Few occurrences" = "#E35932FF", "Exporting interceptions" = "#BB3754FF", "Localized" = "#BB3754FF", "Present" = "#89226AFF", "Widespread" =  "#56106EFF"))+
  scale_color_manual(values = c("Unreported" = "#FCB519FF", "Eradicated" = "#FCB519FF", "No longer present" = "#FCB519FF", "Intercepted only" = "#FCB519FF", "Transient under eradication" = "#F98C0AFF", "Transient under surveillance" = "#F98C0AFF", "Few occurrences" = "#E35932FF", "Exporting interceptions" = "#BB3754FF", "Localized" = "#BB3754FF", "Present" = "#89226AFF", "Widespread" =  "#56106EFF"))+
  coord_fixed()+
  theme_graph()+
  theme(legend.position = "right", 
        legend.key.size = unit(1, 'cm'))

focused.net<-focused.net + 
  guides(fill = guide_legend(title = "Reported distribution \nTuta absoluta",
                                title.position = "top", 
                                override.aes = list(size = 10), 
                             title.theme = element_text(size = 15), 
                             label.theme = element_text(size = 15)),
         color = guide_legend(title = "Reported distribution \nTuta absoluta",
                                title.position = "top", 
                                override.aes = list(size = 10), 
                             title.theme = element_text(size = 15), 
                             label.theme = element_text(size = 15)),
            size = guide_legend(title = "log10( \ntomato \nacreage)", 
                                title.theme = element_text(size = 15), 
                                label.theme = element_text(size = 15)), 
            edge_width = guide_legend(title = "Pest trade \nmovement potential", 
                                      title.theme = element_text(size = 15), 
                                      label.theme = element_text(size = 15))
         )
focused.net
#ggsave(plot = focused.net, filename = "Tuta-absoluta-tomato-net.svg",
#       width = 15, height = 11, units = "in", dpi = 250)
```


## Introduction risk in the United States of America
```{r, fig.width=15, fig.height=10, dpi=200, message=FALSE}
focused.countries<-c("USA")
focused.PestMove<-PestMoveLinks %>%
  filter(ImportCountry %in% focused.countries)
focused.TradingCountries<-TradingCountries %>% 
  filter(ExportCountry %in% unique(c(focused.PestMove$ExportCountry,focused.PestMove$ImportCountry)))
focused.PestNet = graph_from_data_frame(d = focused.PestMove,
                                        vertices = focused.TradingCountries,
                                        directed = T)
V(focused.PestNet)$ExtentCat<-factor(V(focused.PestNet)$ExtentCat,
                                 levels = c("Unreported", "Eradicated", "No longer present", "Intercepted only", "Transient under eradication", "Transient under surveillance", "Few occurrences", "Exporting interceptions", "Localized", "Present", "Widespread"))

focused.net <- ggraph(focused.PestNet, layout = "stress")+
  geom_edge_link(aes(edge_width = GeoTradeRisk, 
                      start_cap = circle(.3), 
                      end_cap = circle(.3)),                  
                  edge_colour = "darkgrey",
                  arrow = arrow(angle = 30, 
                                length = unit(0.15, "inches")))+
  geom_node_point(aes(fill = ExtentCat,
                      color = ExtentCat,
                      size = HostArea), 
                  shape = 21)+
  geom_node_text(aes(label = name), size = 6, 
                 repel = TRUE)+
  scale_edge_width_continuous(range = c(0.1,1.5))+
  scale_size_continuous(range = c(1,10))+
  scale_fill_manual(values = c("Unreported" = "#FCB519FF", "Eradicated" = "#FCB519FF", "No longer present" = "#FCB519FF", "Intercepted only" = "#FCB519FF", "Transient under eradication" = "#F98C0AFF", "Transient under surveillance" = "#F98C0AFF", "Few occurrences" = "#E35932FF", "Exporting interceptions" = "#BB3754FF", "Localized" = "#BB3754FF", "Present" = "#89226AFF", "Widespread" =  "#56106EFF"))+
  scale_color_manual(values = c("Unreported" = "#FCB519FF", "Eradicated" = "#FCB519FF", "No longer present" = "#FCB519FF", "Intercepted only" = "#FCB519FF", "Transient under eradication" = "#F98C0AFF", "Transient under surveillance" = "#F98C0AFF", "Few occurrences" = "#E35932FF", "Exporting interceptions" = "#BB3754FF", "Localized" = "#BB3754FF", "Present" = "#89226AFF", "Widespread" =  "#56106EFF"))+
  coord_fixed()+
  theme_graph()+
  theme(legend.position = "right", 
        legend.key.size = unit(1, 'cm'))

focused.net + 
  guides(fill = guide_legend(title = "Reported distribution \nTuta absoluta",
                                title.position = "top", 
                                override.aes = list(size = 10), 
                             title.theme = element_text(size = 15), 
                             label.theme = element_text(size = 15)),
         color = guide_legend(title = "Reported distribution \nTuta absoluta",
                                title.position = "top", 
                                override.aes = list(size = 10), 
                             title.theme = element_text(size = 15), 
                             label.theme = element_text(size = 15)),
            size = guide_legend(title = "log10( \ntomato \nacreage)", 
                                title.theme = element_text(size = 15), 
                                label.theme = element_text(size = 15)), 
            edge_width = guide_legend(title = "Pest trade \nmovement potential", 
                                      title.theme = element_text(size = 15), 
                                      label.theme = element_text(size = 15))
         )
```


## Pest introduction risk based on multiple network metrics
```{r}
# Calculating node centralities for each trading country
E(pestMoveNet)$weight <- E(pestMoveNet)$GeoTradeRisk
country.strength <- strength(pestMoveNet, v = V(pestMoveNet), mode = c("in"), weights = E(pestMoveNet)$weight)

trans.weight<-1.0001*max(E(pestMoveNet)$GeoTradeRisk) - E(pestMoveNet)$GeoTradeRisk
country.betweenness<-igraph::betweenness(pestMoveNet,v=V(pestMoveNet), directed = TRUE, weights = trans.weight)

#country.PageRank <- page_rank(pestMoveNet, vids = V(pestMoveNet), directed = TRUE, weights = E(pestMoveNet)$GeoTradeRisk)
#PageRank can be used better to measure a node's influence on other nodes as one component of PageRank scores is the number of outgoing links from a node (https://doi.org/10.1016/S0169-7552(98)00110-X).

country.eigenvector <- eigen_centrality(pestMoveNet, directed = TRUE, weights = E(pestMoveNet)$GeoTradeRisk, scale = TRUE)$vector

country.centralities<-data.frame(country.strength, country.betweenness, 
                                 country.eigenvector$vector)
country.centralities<-data.frame(country.strength, country.betweenness, 
                                 country.eigenvector, joint.risk)
country.centralities<-country.centralities%>%
  mutate(country.strength=country.strength/max(country.strength),
         country.betweenness=country.betweenness/max(country.betweenness),
         PTRI = country.strength+country.betweenness+country.eigenvector+joint.risk,
         PTRI = PTRI / max(PTRI))
```


## Pest introduction risk based on multiple network metrics
```{r, fig.width=18, fig.height=15, dpi=300}
multi.net <- ggraph(pestMoveNet, layout = "centrality", cent = country.centralities$PTRI^(1/2))+
  geom_edge_link(aes(edge_width = GeoTradeRisk, 
                      start_cap = circle(.3), 
                      end_cap = circle(.3)),                  
                  edge_colour = "darkgrey",
                  arrow = arrow(angle = 30, 
                                length = unit(0.15, "inches")))+
  geom_node_point(aes(fill = ExtentCat,
                      color = ExtentCat,
                      size = exp(HostArea),
                      shape = country.type))+
  geom_node_text(aes(label = name), size = 6, 
                 repel = TRUE)+
  scale_shape_manual(values = c("Intermediary"=23, "Producer"=21))+
  scale_edge_width_continuous(range = c(0.05,2))+
  scale_size_continuous(range = c(2,12))+
  scale_fill_manual(values = c("Unreported" = "#FCB519FF", "Eradicated" = "#FCB519FF", "No longer present" = "#FCB519FF", "Intercepted only" = "#FCB519FF", "Transient under eradication" = "#F98C0AFF", "Transient under surveillance" = "#F98C0AFF", "Few occurrences" = "#E35932FF", "Exporting interceptions" = "#BB3754FF", "Localized" = "#BB3754FF", "Present" = "#89226AFF", "Widespread" =  "#56106EFF"))+
  scale_color_manual(values = c("Unreported" = "#FCB519FF", "Eradicated" = "#FCB519FF", "No longer present" = "#FCB519FF", "Intercepted only" = "#FCB519FF", "Transient under eradication" = "#F98C0AFF", "Transient under surveillance" = "#F98C0AFF", "Few occurrences" = "#E35932FF", "Exporting interceptions" = "#BB3754FF", "Localized" = "#BB3754FF", "Present" = "#89226AFF", "Widespread" =  "#56106EFF"))+
  coord_fixed()+
  theme_graph()+
  theme(legend.position = "right", 
        legend.key.size = unit(1, 'cm'))

multi.net + 
  guides(fill = guide_legend(title = "Reported distribution: \nTuta absoluta",
                                title.position = "top", 
                                override.aes = list(size = 10), 
                             title.theme = element_text(size = 15), 
                             label.theme = element_text(size = 15)),
         color = guide_legend(title = "Reported distribution: \nTuta absoluta",
                                title.position = "top", 
                                override.aes = list(size = 10), 
                             title.theme = element_text(size = 15), 
                             label.theme = element_text(size = 15)),
         shape = guide_legend(title = "Country category",
                                title.position = "top", 
                                override.aes = list(size = 10), 
                             title.theme = element_text(size = 15), 
                             label.theme = element_text(size = 15)),
         size = guide_legend(title = "log10( \ntomato \nacreage)", 
                                title.theme = element_text(size = 15), 
                                label.theme = element_text(size = 15)),
         edge_width = guide_legend(title = "Pest trade \nmovement potential", 
                                      title.theme = element_text(size = 15), 
                                      label.theme = element_text(size = 15))
         )
```



